Molecular dynamics simulation of synchronization of a driven particle

### 2021-04-13 manuscript/manuscript_ajp_version3.tex
fixed intro - todo conclusion

\begin{abstract}
  We discuss a 
  molecular dynamics simulation
  of a single particle
  moving through
  a viscous liquid
  while being driven 
  across a washboard potential energy landscape.
  Our results show many dynamical patterns
  as the landscape and driving force are altered.
  For certain conditions,
  the particle velocity and location
  are synchronized or 
  phase-locked,
  forming closed orbits in phase space.
  Quasi-periodic motion is common, 
  for which the
  dynamical center of motion shifts the
  phase space orbit.
  %DM removed first sentence,
  %which was strong and clear, but isolated in the abstract
  %put those ideas in this final sentence, making it a mouthful [26 words]
  Isolating
  synchronized motion
  in 
  simulations and table-top experiments,
  can be used to study %DM replaced isolate with study
  complex natural behaviors
  which
  play an important role in many physical processes.
 %[suggest omitting this sentence. Code not in text. We include molecular dynamics code
 % to simulate and characterize the
  %particle dynamics.] 
\end{abstract}

@@

\section{Introduction} 
Synchronization is a universal phenomena
in which
individual oscillators change frequency due
to external stimuli.\cite{Pikovsky2003}
The
flickering patterns of
candle flames 
mediated by temperature fluctuations,\cite{Okamoto2016}
the vibrations of singing wine glasses interacting 
through sound waves,\cite{Arane2009}
and metronomes vibrating through a supporting platform\cite{Jia2015}
are examples of in-phase coupled oscillations. 
Biological systems benefit from cooperative
synchronization --
birds  coordinate wing flaps
to optimize energy use during flight,\cite{Portugal2014}
frogs alternate croaking patterns,\cite{Aihara2014}
humans clap in time with music,\cite{Tranchant2016}
and 
at a cellular level, 
neurons simultaneously fire in cardiac muscle\cite{MartinHall1999}
and brain tissue.\cite{Singer1999}
%DM ended paragraph with fewer examples, 
%Environmental conditions
%DM add add transition sentence
%
External forcing
can cause or regulate synchronization.
%where external forces
%add or remove energy from the system.
For example, 
an electrical pacemaker 
regulates a heart beat and
a pulsed light modifies the
flashing pattern of fireflies.

The model discussed in this paper
resembles an externally forced pendulum.
%damping
%transition
%%DM - focus more on single particle with environment
A single pendulum 
may swing in time with an %be driven or damped by
driving force
like  %a skateboarder in a half-pipe or
a pushed child on a swing,
and may be damped
by mechanical friction.
%of gravity to return
%to an equilibrium position,
%or  to decreasing amplitude
% or air resistance.
Synchronized phase-locking or mode-locking 
first appeared in the scientific literature
in 1665 with 
Huygens'  experiments on the
motions of synchronized 
pendula in wall-mounted clocks.\cite{Bennett2002}
Clock pendula swing in phase
when coupled through shared vibrations.
A locked-mode is an integer frequency ratio.
In Huygens clocks,
the pendula were observed to swing at the same rate
in the same direction (1:1 mode) or
opposite directions ($-1$:1 mode).
A 2:1 mode occurs when a simple pendulum four
times the length of another swings with twice the period.
%DM transition sentence
Many subsequent 
simulations and laboratory experiments
have revealed 
mode-locked synchronization
in a diverse range of 
oscillation phenomena.

%Single oscillators 
%can demonstrate a variety of synchronized behaviors.
Colloid particles in complex environments are an excellent
model system to explore %these 
the dynamics of mode-locking. %synchronization.
In experiments %DM moved the ``in experiments'' to second sentence
typical colloids are   
plastic spheres suspended in
de-ionized water or silica beads suspended in organic solvent.
Colloids can be trapped
with 
radiation pressure from 
a laser beam.\cite{Ashkin1997} 
A colloid in such a confining region,
an optical trap, oscillates
about a equilibrium point.
A colloid centered in an optical trap is 
uniformly bombarded by photons. 
Off-center colloids 
experience a net force
due to uneven photon collisions across
the particle surface.
Depending on the 
location of the particle in the trap,
the radiation pressure either moves colloids toward the center 
or ejects it from the trap.
Diffraction gratings can create
more complex light environments, 
such as periodic patterns of minima
suitable for synchronization studies.\cite{Grier2003}
A driven colloid
can demonstrate
synchronized
motion within one trap,
or synchronized hopping between multiple traps.

Colloids have been
shown to 
provide an ideal system to 
control and study
phase-locked motions
%are useful for
%understanding synchronization
at a single particle level.\cite{Juniper2015,Juniper2017}
% such as synchronized mode-locking can be studied with
%An advantage of colloid studies
While many systems oscillate,
trapped colloids are useful for 
providing controlled 
measurements of the
step-by-step dynamics of particle motion.
Since
colloids are large 
and move slowly, 
%colloids are considered model systems
particle positions 
can be measured in real time 
with an optical camera.\cite{Pertsinidis2001}
%Light is a
%tool for manipulating the colloidal environment,
%which can alter synchronization patterns.
%Synchronization patterns can be altered
%using light to manipulate
%the colloidal environment.
Simulations provide complementary
information,
where environmental and driving parameters may be precisely
and quickly modified to explore
the many oscillation patterns.
%the dynamics from external
In the following study,
we confine colloids  
to move along a periodic light pattern
with minima and maxima defined by a sine wave,
a washboard potential energy substrate.

A colloid driven across a periodic
substrate may 
synchronize its location
to the pattern of the external drive.
%moving back and forth in time with
%the beat, or
%moving between substrate minima.
When 
a constant or 
dc drive is applied,
the particle velocity is modulated by 
the potential energy landscape exerted on the particle.   
Below some threshold  
the dc force is not strong enough to push the particle
across a potential maximum so the average velocity is zero,
a phenomena referred to as pinning.\cite{Reichhardt2017}
Above the pinning threshold,
a dc driven particle
%a particle subject to a constant
%drive force
increases its speed at a rate proportional
to the external drive.  
When the applied force varies periodically,
an ac drive,
%cause
the particle can hop back and forth across
the landscape minima.
Many synchronized patterns occur,
controlled by the substrate pattern and driving force.
When a particle is driven by combined 
ac and dc forces,
mode-locking appears 
where 
the average particle velocity
%of an  driven particle 
is fixed for a range of dc drive forces.\cite{Reichhardt2015}

@@

\section{Simulation}
\label{sec:MD}
We use a classical model for 
studying the dynamics, 
using the net force on a particle to calculate
its trajectory.
The particle is confined in a two-dimensional 
simulation of area $A = L \times L$, where $L=46.\bar{6}\,a_0$
and $a_0$ is a dimensionless unit of length. (In the simulations we   set $a_0 = 1$.)
The particle 
has
position $\vec{r} = x \hat{x} + y \hat{y}$
and velocity $\vec{v} = d\vec{r}/dt$.
The edges of the system are treated by
periodic boundary conditions,
such that a particle leaving the edge of the system is mapped
back to a position within the simulation boundaries 
by the transformation $x+L \rightarrow x$ and $y+L \rightarrow y$.
We show a schematic of the system in Fig.~\ref{fig:1_landscape}(a).
The units of the simulated variables are summarized in Table~\ref{tab:1}.

@@

\section{Conclusion}
\label{sec:conclusion}	
A single particle driven across a periodic potential landscape 
synchronizes its motion 
to environmental and external forces.  
Our simulations reproduce the experiments and simulations presented in 
Juniper {\it et al.} \cite{Juniper2015, Juniper2017}
of 
mode locking in
driven colloids on a
periodic optical landscape.
The model
is easy to simulate yet relevant
to a broad range of condensed matter systems.
Colloids are 
relatively easy to 
manipulate and image in experiments,
making ideal proxies 
for systems 
such as cold atoms or electron gases.\cite{Grier2003}
Dynamical mode-locking 
is 
observed in  technologies such as 
 quantum electronic
devices as 
stepped regions in the relation between current and voltage,
where the voltage is the analog of the external driving force
and current is the analog of particle velocity.
%DM FIX FOLLOWING SENTENCE
Known as Shapiro steps, 
these mode-locked or phase-locked currents
%arise 
%due to applied ac voltages
were first
observed in single Josephson junctions\cite{Shapiro1963, Golubov2004}.
%coupled arrays of junctions.\cite{Benz1990}
%DM END FIX
Shapiro steps vary in width depending on the strength of the
applied ac forces,
and are observed in a variety of ac and dc driven systems
displaying
non-Ohmic behavior in voltage-current curves,
including
charge density waves, spin density waves
and superconducting vortices in landscapes 
engineered with periodic patterns of pinning sites.\cite{Reichhardt2000}

Mode-locking is a useful probe 
of complex quantum mechanical systems
because the motions of individual particles can  be inferred only
from other measurements.
Our results are relevant 
to synchronization effects
in a broad range of experimental systems
including optically confined colloids,
superconductors with periodic pinning arrays, 
and the charge and spin of atomic systems.

>>>>

\begin{abstract}
  Synchronization
  plays an important role in many physical processes.
  We discuss synchronization in a 
  molecular dynamics simulation
  of a single particle
  moving through
  a viscous liquid
  while being driven 
  across a washboard potential energy landscape.
  Our results show many dynamical patterns
  as the landscape and driving force are altered.
  For certain conditions,
  the particle velocity and location
  are synchronized or 
  phase-locked,
  forming closed orbits in phase space.
  Quasi-periodic motion is common, 
  for which the
  dynamical center of motion shifts the
  phase space orbit.
  %DM removed first sentence,
  %which was strong and clear, but isolated in the abstract
  %put those ideas in this final sentence, making it a mouthful [26 words]
  By 
  isolating
  synchronized motion
  in 
  simulations and table-top experiments,
  we 
  can study %DM replaced isolate with study
  complex natural behaviors
  important to many physical processes.
 %[suggest omitting this sentence. Code not in text. We include molecular dynamics code
 % to simulate and characterize the
  %particle dynamics.] 
\end{abstract}

@@

\section{Introduction} 
Synchronization is a universal phenomena
in which
individual oscillators change frequency due
to external stimuli.\cite{Pikovsky2003}
The
flickering patterns of
candle flames 
mediated by temperature fluctuations,\cite{Okamoto2016}
the vibrations of singing wine glasses interacting 
through sound waves,\cite{Arane2009}
and metronomes vibrating through a supporting platform\cite{Jia2015}
are examples of in-phase environmentally coupled oscillations. 
Biological systems benefit from cooperative
synchronization --
birds  coordinate wing flaps
to optimize energy use during flight,\cite{Portugal2014}
frogs alternate croaking patterns,\cite{Aihara2014}
humans clap in time with music,\cite{Tranchant2016}
and 
at a cellular level, 
neurons simultaneously fire in cardiac muscle\cite{MartinHall1999}
and brain tissue.\cite{Singer1999}
%DM ended paragraph with fewer examples, 
%Environmental conditions
%DM add add transition sentence
%
External forcing
can cause or regulate synchronization.
%where external forces
%add or remove energy from the system.
For example, 
an electrical pacemaker 
regulates a heart beat and
a pulsed light modifies the
flashing pattern of fireflies.

The model discussed in this paper
resembles an externally forced pendulum.
%damping
%transition
%%DM - focus more on single particle with environment
A single pendulum 
can swing in time with an %be driven or damped by
driving force
like  %a skateboarder in a half-pipe or
a pushed child on a swing,
and may be damped
by mechanical friction.
%of gravity to return
%to an equilibrium position,
%or  to decreasing amplitude
% or air resistance.
We focus on 
synchronized phase-locking or mode-locking,
which 
first appeared in the scientific literature
in 1665 with 
Huygens'  experiments on the
motions of %synchronized 
pendula in wall-mounted clocks.\cite{Bennett2002}
A locked-mode is an integer frequency ratio
in coupled oscillators. %Clock pendula swing in phase
%when coupled through shared vibrations.
In Huygens' clocks,
the mode-locked pendula were observed to swing at the same rate
in the same direction (1:1 mode) or
opposite directions ($-1$:1 mode).
Higher modes might be achieved
by altering the pendula lengths - 
a 2:1 mode occurs when a simple pendulum
swings with twice the period
of another one quarter the length.
%DM transition sentence
%Many subsequent 
Simulations and laboratory experiments
have revealed 
mode-locked synchronization
in a diverse range of 
oscillation phenomena.

%Single oscillators 
%can demonstrate a variety of synchronized behaviors.
Colloid particles in complex environments are an excellent
model system to explore %these 
the dynamics of mode-locking. %synchronization.
In experiments %DM moved the ``in experiments'' to second sentence
typical colloids are   
plastic spheres suspended in
de-ionized water or silica beads suspended in organic solvent.
Colloids can be trapped
with 
radiation pressure from 
a laser beam.\cite{Ashkin1997} 
A colloid in such a confining region,
an optical trap, oscillates
about a equilibrium point.
A colloid centered in an optical trap is 
uniformly bombarded by photons. 
Off-center colloids 
experience a net force
due to uneven photon collisions across
the particle surface.
Depending on the 
location of the particle in the trap,
the radiation pressure either moves colloids toward the center 
or ejects it from the trap.
Diffraction gratings can create
more complex light environments, 
such as periodic patterns of minima
suitable for synchronization studies.\cite{Grier2003}

Colloids have been
shown to 
provide an ideal system to 
control and study
phase-locked motions
%are useful for
%understanding synchronization
at a single particle level.\cite{Juniper2015,Juniper2017}
% such as synchronized mode-locking can be studied with
%An advantage of colloid studies
A driven colloid
can demonstrate
synchronized
motion within one trap,
or synchronized hopping between multiple traps. 
While many systems oscillate,
trapped colloids are useful for 
providing controlled 
measurements of the
step-by-step dynamics of particle motion.
Since
colloids are large 
and move slowly, 
%colloids are considered model systems
particle positions 
can be measured in real time 
with an optical camera.\cite{Pertsinidis2001}
%Light is a
%tool for manipulating the colloidal environment,
%which can alter synchronization patterns.
%Synchronization patterns can be altered
%using light to manipulate
%the colloidal environment.
Simulations provide complementary
information,
where environmental and driving parameters may be precisely
and quickly modified to explore
the many oscillation patterns.
%the dynamics from external
In the following study,
we confine colloids  
to move along a periodic light pattern
with minima and maxima defined by a sine wave,
a washboard potential energy substrate.

A colloid driven across a periodic
substrate may 
synchronize its location
to the pattern of the external drive.
%moving back and forth in time with
%the beat, or
%moving between substrate minima.
When 
a constant or 
dc drive is applied,
the particle velocity is modulated by 
the potential energy landscape exerted on the particle.   
Below some threshold  
the dc force is not strong enough to push the particle
across a potential maximum so the average velocity is zero,
a phenomena referred to as pinning.\cite{Reichhardt2017}
Above the pinning threshold,
a dc driven particle
%a particle subject to a constant
%drive force
increases its speed at a rate proportional
to the external drive.  
When the applied force varies periodically,
an ac drive,
%cause
the particle can hop back and forth across
the landscape minima.
When a particle is driven across a substrate
by combined 
ac and dc forces,
it is possible for the particle to demonstrate 
mode-locking.
In these modes,  
%Due to the combination of external drive
%and substrate minima,
the particle hops at a controlled rate such that 
the average particle velocity is 
%of an  driven particle 
fixed for a range of dc drive forces
with a magnitude proportional to the
landscape period and driving frequency.\cite{Reichhardt2015}
%Many synchronized patterns occur,
%controlled by the substrate pattern and driving force.

@@

\section{Simulation}
\label{sec:MD}
We use a classical model for 
studying the dynamics, 
using the net force on a particle to calculate
its trajectory.
The particle is confined in a one-dimensional %DM reduced to 1D
simulation of length $L=46.\bar{6}\,a_0$, %DM 1D
where $a_0$ is a dimensionless unit of length.
(In the simulations we set $a_0 = 1$.)
The particle 
has
position $\vec{r} = y \hat{y}$ %DM 1D
and velocity $\vec{v} = d\vec{r}/dt$.
The edges of the system are treated by
periodic boundary conditions,
such that a particle leaving the edge of the system is mapped
back to a position within the simulation boundaries 
by the transformation $y+L \rightarrow y$. %DM 1D
We show a schematic of the system in Fig.~\ref{fig:1_landscape}(a).
The units of the simulated variables are summarized in Table~\ref{tab:1}.

@@

\section{Conclusion}
\label{sec:conclusion}

Our simulations reproduce the experiments and simulations presented in 
Juniper {\it et al.} \cite{Juniper2015, Juniper2017}
of 
mode locking in
driven colloids on a
periodic optical landscape.
A single particle driven across a periodic potential landscape 
synchronizes its motion 
to environmental and external forces. 
%The model
%is easy to simulate yet relevant
%to a broad range of condensed matter systems.
Our results are relevant 
to synchronization effects
in a broad range of experimental systems
including optically confined colloids,
superconductors with periodic pinning arrays, 
and the charge and spin of atomic systems.

Colloids are 
relatively easy to 
manipulate and image in experiments,
making ideal proxies 
for systems 
such as cold atoms or electron gases.\cite{Grier2003}
%DM FIX FOLLOWING SENTENCE
Dynamical mode-locking 
is 
observed in technologies such as 
quantum electronic
devices as 
stepped regions in the relation between current and voltage,
where the voltage is the analog of the external driving force
and current is the analog of particle velocity.
%DM FIX FOLLOWING SENTENCE
Known as Shapiro steps, 
these mode-locked or phase-locked currents
%arise 
%due to applied ac voltages
were first
observed in single Josephson junctions\cite{Shapiro1963, Golubov2004}.
%coupled arrays of junctions.\cite{Benz1990}
%DM END FIX
Shapiro steps vary in width depending on the strength of the
applied ac forces,
and are observed in a variety of ac and dc driven systems
displaying
non-Ohmic behavior in voltage-current curves,
including
charge density waves, spin density waves
and superconducting vortices in landscapes 
engineered with periodic patterns of pinning sites.\cite{Reichhardt2000}
Mode-locking is a useful probe 
of complex quantum mechanical systems
because the motions of individual particles can  be inferred only
from other measurements.
%insert final sentence summarizing all the things.
